### 2009-09-20 latex/paper1/contents.tex
paper

\newcommand{\din}[0]{\ensuremath{\delta_{in}}}
\newcommand{\dout}[0]{\ensuremath{\delta_{out}}}

% TODO: say that we use DIRECTED GRAPHS

% DEPENDENCY ANALYSIS
%
% component dependency isn't a good description...
% reference?
% usage?

% Martin Fowler
%
%   Coupling also occurs when code in one
%module uses code from another, perhaps by
%calling a function or accessing some data. 
%   When I use the term dependency, I use it as
%defined in the Unified Modeling Language
%(UML). So, the UI module depends on the do-
%main module if any code in the UI module ref-
%erences any code in the domain model - by
%calling a function, using some data, 
%or using types defined in the domain module
%
% A dependency is unidirectional

%A dependency is a semantic relationship where a change to the influent or
%independent modeling element may affect the semantics of the dependent
%modeling element. http://en.wikipedia.org/wiki/Dependency_(UML)

% Neeraj Sangal
% Dependencies are extracted from the code by a conventional static analysis
%
% Ao falar de dependências, cita
%   Designing Software for Ease of Extension and Contraction
%
%The LDM tool uses a standard notion of dependency, in which a
%module A depends on a module B if there are explicit references
%in A to syntactic elements of B. Currently, a module is a Java
%class, but a more fine-grained analysis is possible. This simple but
%effective notion of dependency works well for understanding
%design dependencies, in which modifications to one module might
%affect another. It is less well suited to determining runtime
%properties (such as how failures can propagate between modules),
%which require a deeper static analysis.
%
%---------
%
% Parnas
%
% A uses B if correct execution of B may be necessary for A to complete the
% task described in its specification.
%
% A uses B if there exist situations in which the correct functioning of A
% depends upon the availability of a correct implementation of B.
%
% "Uses" can also be formulated as "requires the presence of a correct version
% of".
%
% Uses vs. invokes

\section{Abstract}

Software clustering algorithms automatically recognize a system's modular
structure by analyzing its implementation.  Due to the lack of well document
software systems, though, the issue of testing these algorithms is still
underexplored, limiting both their adoption in the industry and the development
of better algorithms.  We propose to rely on software models to produce
arbitrarily large test sets. In this paper we consider three such models and
analyze how similar the artifacts they produce are from artifacts from real
software systems.

\section{Introduction} \label{sec:introduction}

%   I want less time, so I hire more programmers
%   Thee project can consume more time, due to comunication costs between
% programmers.
%   Unless there is a good division of tasks
%
% (needn't be large)

% TODO: this paragraph can mostly disappear.
%
The time spent on the maintenance of a software system can be affected by the
assignment of developers to implementation components of the system.
If two development teams are responsible for maintaining components that are
highly interdependent, they need to coordinate their work often in order avoid
code duplication and breaking each other's component funcionality.

%One of the most obvious ways to reduce the time needed to develop or maintain a
%software system is to assign more developers to it. When combined with poor
%management, though, adding developers may in fact raise the development time,
%due to the increasing communication costs between developers \cite{Brooks1995}.
%
%The communication costs are affected by the assignment of developers to
%implementation components of a system.
%%The assignment of developers to implementation components of a system affects
%%the amount of communication needed between developers. 
%If two development teams are responsible for maintaining components that are
%highly interdependent, they need to coordinate their work often in order avoid
%code duplication and breaking each other's component funcionality.

Therefore, to optimize the development, a development manager can decompose the
system into modules, or groups of components, so that there are as few
inter-module dependencies as possible.% (i.e., the modules are weakly couple.).
Each module is, then, threated as an unit of work assignment \cite{Parnas1972}:
each module is assigned to one team and, because of the little inter-module
dependency, the communication between teams is reduced, therefore speeding up
the development.

Therefore, to decompose a system, a manager needs to have global knowledge of
the dependencies between its components. Although dependency analysis tools can
extract such information from the source code of the system, the task of finding
modules can be overwhelming because of the large number of components and
dependencies.

Software clustering algorithms, also known as architecture recovery algorithms
\cite{Mancoridis1998,Anquetil1999,Tzerpos2000,Andritsos2005},
%\cite{Pollet2007}, 
automatically decompose a software system modules by analyzing the network of
dependencies between its components.  Although many algorithms have been
proposed, there are few comparative studies.
%empirical studies regarding the quality of the decompositions they find.
%EVALUATION CRITERION
One important test for a software clustering algorithm is to apply it to a
collection of software systems with known reference decompositions made by
experienced developers \cite{Anquetil1999}. It is expected that a good algorithm
finds decompositions that are similar to reference decompositions, which can be
measured by a similarity metric such as MoJo \cite{Tzerpos1999} or EdgeSim
\cite{Mitchell2001}.

Unfortunately there are few publicly available reference decompositions
\cite{Koschke2000} and, as a result, few comparative studies exist. Also,
because it is costly to obtain reference decompositions for large systems, most
studies test the algorithms with a couple of small and medium systems
\cite{Anquetil1999,Maqbool2007,Bittencourt2009}.

In our research we propose to evaluate software clustering algorithms by
applying them to synthetic, i.e., computer generated, component dependency
networks with built-in reference decompositions. This approach enables one to
test algorithms with arbitrarily large samples in a controlled manner.

Of course, it is desirable that the synthetic networks resemble networks
extracted from real software systems. In this paper, we present three network
models and show empirically that all three models are capable of synthesizing
networks that resemble software networks. We consider object-oriented systems
with classes as components.

The paper is organized as follows:
Section \ref{sec:models} presents three network models.
%
Section \ref{sec:characterization} demonstrates a method to tell if a network
resembles a software network.
%
Section \ref{sec:evaluation} describes an experiment that shows that the three
models synthesize networks that resemble software networks.
%
Section \ref{sec:conclusion} discusses the impact of this research and future
work.

%  \begin{figure*}[!t]
%  \centering
%  \includegraphics[width=1.0\textwidth]{diagram}
%  \caption{Evaluation of a software clustering recovery algorithm}
%  \label{fig:diagram}
%  \end{figure*}

\section{Network Models} \label{sec:models}

Network theory studies general properties of many types of networks by using
statistical analysis. In the last decade, it has been found that many networks
arising from sociology, biology, technology and other domains present remarkable
structural similarities. It has been shown that, in these networks, the
distribution of vertex degrees is a power law, i.e., the number of vertices
connected to $k$ edges is proportional to $k^{-\gamma}$, where $\gamma$ is a
positive constant. These networks are called scale-free networks
\cite{Barabasi1999}.

Network theory has been applied to class dependency networks and it was shown
that they are also scale-free networks \cite{Myers2003,Valverde2002}. Thus, the
study of the structure of object-oriented software systems can benefit from the
vast literature available about scale-free networks.

In particular, many scale-free network models were proposed. These models are
algorithms that were proven, either formally or empirically, to synthesize
networks that are scale-free. In this section we present three models that
synthesize scale-free networks with built-in modular decompositions: BCR+, CGW,
and LF. Their output can be used to test software clustering algorithms.

% directed unweighted graphs

%\subsection{ER (Erdos-Renyi)}
%\subsection{Configuration model} %XXX consider using with the most typical sw

\subsection{BCR+}

The BCR model \cite{Bollobas2003} aims to model the network of hyperlinks
between web pages as a directed graph without modules. We have developed an
extension to this model, called BCR+, that adds modules to the construction of
the network. The model accepts the following parameters:

\begin{itemize}
\item number of vertices, $n$;
\item a directed graph of modules, $G$;
\item three probabilities, $p_1$, $p_2$, and $p_3$, summing to 1;
\item constant $\mu$, with $0 \le \mu \le 1$;
\item base in-degree, $\din$;
\item base out-degree, $\dout$.
\end{itemize}

The graph $G$ contains one vertex for each module that will be created and
determines a ``supply'' relationship between modules. We say that module $M_2$
is a supplier of module $M_1$ if, and only if, $G$ contains an edge from $M_1$
to $M_2$.

Networks generated by the model can contain both internal edges and external
edges. An internal edge connects two vertices in the same module. An external
edge connects vertices in distinct modules. In this model, an external edge from
a vertex $v_1 \in M_1$ to a vertex $v_2 \in M_2$ is only allowed if $M_2$ is a
supplier of $M_1$.

The parameter $\mu$ controls the proportion external edges in the network. Lower
values of $\mu$ lead to networks with fewer external edges and, therefore, to
weakly coupled modules.

First the model creates an initial network in which each module contains exactly
one vertex. Then all allowed external edges are added. Finally, the networks is
modified by sucessive choices between three operations that add vertices or
edges to the network, until it reaches $n$ vertices. Each time, the probability
of choosing the $i$-th operation is $p_i$.

Before describing the operations, some definitions are needed. The expression
in-degree($x$) means the number of edges that enter the vertex $x$;
out-degree($x$), likewise, is the number of edges that leave vertex $x$. The
expression ``to choose a vertex according to f(x)'' means that the probability
of choosing a vertex $x$ is given by the following probability function:

$$
  \mathrm{P}(x) ~=~ \frac{ \mathrm{f}(x) }
  { \displaystyle\sum_{i} \mathrm{f}(i) }
$$

(the denominator is a normalizing factor, so the sum of probabilities is 1).

The three operations are described next:

\begin{enumerate}

\item \emph{Adding a vertex with an outgoing edge}. An existing vertex, $w$, is
chosen according to $\mathrm{f}(x) = \din + \mathrm{in\mbox{-}degree}(x)$. A new
vertex, $v$, is added to the module of $w$, together with an edge from $v$ to
$w$.

\item \emph{Adding a vertex with an ingoing edge}. An existing vertex, $w$, is
chosen according to $\mathrm{f}(x) = \dout + \mathrm{out\mbox{-}degree}(x)$. A
new vertex, $v$, is added to the module of $w$, together with an edge from $w$
to $v$.

\item \emph{Adding an edge between existing vertices}. A vertex, $v$, is choosen
from the network according to $\mathrm{f}(x) = \dout +
\mathrm{out\mbox{-}degree}(x)$.  Then an edge is added from vertex $v$ to an
existing vertex $w$, which is chosen according to one of the following cases:

\begin{enumerate}
  \item with probability $\mu$, $w$ is chosen among vertices in modules that are
  suppliers of the module of $v$;
  \item with probability $1 - \mu$, $w$ is chosen among vertices in that same
  module as $v$.
\end{enumerate}

\end{enumerate}

It is easy to see that nodes with high in-degree are more likely to receive new
ingoing edges. The parameter $\din$ can reduce this bias by providing a base
in-degree that is applied to all vertices when computing the probabilities.
Consider two vertices, $v_1$ with in-degree 4, and $v_2$ with in-degree 8. If
$\din = 0$, $v_2$ is twice more likely to receive a new incoming edge; if,
otherwise, $\din = 4$, $v_2$ is only $\frac{3}{2}$ more likely to receive the
edge. The same reasoning applies to $\dout$.

The BCR+ model is a growth model, meaning that the network is created vertex by
vertex, growing from an initial network. It can, therefore, simulate the
evolution of a software network. Moreover, it can simulate the evolution of a
software system subject to constraints in module interaction, as is the case
with top-down system design methodologies.

%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*

\subsection{CGW}

The CGW model \cite{Chen2008} was proposed to model the evolution of software
systems organized in modules. It accepts the following parameters:

\begin{itemize}
\item number of vertices, $n$;
\item number of modules, $m$;
\item probabilities $p_1, p_2, p_3, p_4$, summing to 1;
\item natural numbers $e_1, e_2, e_3, e_4$;
\item constant $\alpha$, with $\alpha \ge -1$.
\end{itemize}

Because the model is described in detail in the original paper, only an
intuitive notion of the meaning of the parameters is provided here. This model
is similar to BCR+ in which it grows an initial network by sucessive
applications of operations. In this case, there four operations, which are
chosen according to probabilities $p_1$, $p_2$, $p_3$, and $p_4$:

\begin{enumerate}
\item adding a vertex with $e_1$ outgoing edges; 
\item adding $e_2$ edges;
\item rewiring $e_3$ edges;
\item removing $e_4$ randomly chosen edges.
\end{enumerate}

The constant $\alpha$ controls the proportion of edges that connect vertices
from distinct modules. For $\alpha = -1$, most edges will connect distinct
modules. For $\alpha > 0$, most edges will connect vertices in the same module,
and the greater the $\alpha$, more strong is the tendency. For $\alpha = 0$
there is no bias, the two kinds of edges are equaly likely.

%Unlike BCR+, this model does not allow constraints on the connection between
%modules, however it accounts for the rewiring and the removal of edges.

%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*%*

\subsection{LF}

The LF model \cite{Lancichinetti2009} is a very flexible model that can
generated weighted directed networks with overlapping modules, that is, in which
a vertex can belong to more than one module. Unlike the previous models, this is
not a growth model: all vertices are generated at once and then the edges are
added.

There is also a special version of the model in which the edges weights are
discarded and the modules are non overlapping. We used the original
implementation of this version, available at
\url{http://santo.fortunato.googlepages.com/inthepress2}. It accepts the
following parameters:

\begin{itemize}
\item number of nodes, $n$;
\item average in-degree, $k$, with $k < n$;
\item maximum in-degree, $max_k$, with $k \le max_k < n$;
\item mixing parameter, $\mu$, with $0 \le \mu \le 1$;
\item minus exponent for the degree distribution, $\gamma$;
\item minus exponent for the module size distribution, $\beta$;
\item minimum module size, $min_m$;
\item maximum module size, $max_m$, with $max_m \ge min_m$;
\end{itemize}

The sizes of the modules are selected from a power law distribution with
exponent $-\beta$. The mixing parameter, $\mu$, is the proportion of edges in the
network that connect vertices from distinct modules. In this model, some
combinations of parameter values are unfeasible. For example, if $n = 100$ then
$min_m$ cannot be 60 (otherwise there would be modules smaller than the minimum
module size).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Characterization of Software Networks} \label{sec:characterization}

Our research hypothesis is that at least one of the presented models can
synthesize networks that are similar to software networks. A central issue,
thus, is how to measure similarity between networks.

We already know that the models synthesize networks that, just like software
networks, are scale-free. This single property, though, isn't enough to prove
the hypothesis, since many known scale-free networks are not software networks
(e.g., biologic networks). Therefore, the similarity metric should be able to
distinguish between software and non-software networks.

In this section we present a similarity metric and validate the metric by
appplying it to a data set containing both software and non-software networks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Similarity Between Networks}

In a recent work, Milo et al. \cite{Milo2002} proposed to characterize networks
by analyzing their triad concentration. A triad is a network with three vertices
in which all vertices are connected. There are only 13 distinct triads, one for
each possible configuration of directed edges, as shown in Figure
\ref{fig:triads}.

\begin{figure*}[!t]
\centering
\includegraphics[width=1.0\textwidth]{triads}
\caption{The 13 network triads.}
\label{fig:triads}
\end{figure*}

By counting how many times each triad appears in a network, one can build a
triad concentration profile (TCP), which is a vector with 13 numbers that
summarize the local structure of the network. Figure \ref{fig:profiles} shows
the TCP for networks from two distinct domains.

% Word adjacency network: word Y follows word X => X->Y
\begin{figure}[!t]
\center
\includegraphics{tcp}
\caption{Triad concentration profiles (TCP) for two networks. On the left,
network extracted from the software system JabRef (see the Appendix). On the
right, word adjacency network for the Japanese language \cite{Milo2004}.}
\label{fig:profiles}
\end{figure}

Following the work by Milo et al. \cite{Milo2004}, similarity between two
networks can be measured by computing Pearson's correlation coefficient between
the corresponding TCPs, which yields a value between -1 (most dissimilar) and 1
(most similar):

$$
\mathrm{similarity}(a, b) ~=~ 
  \mathrm{cor}(\mathrm{TCP}(a), \mathrm{TCP}(b))\mathrm{,}
$$

where $a$ and $b$ are networks, TCP($x$) is the triad concentration profile for
network $x$, and cor($x$, $y$) is Pearson's correlation coefficient.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Data Set}

To support the evaluation of the metric, we have collected 131 networks. The
networks are described in detail in the Appendix.

\subsubsection{Software networks} We have collected 65 software systems written in
Java, with size ranging from 111 to 35,363 classes. Java was chosen for being a
popular programming language in which many open source systems have been
written. The software networks, representing dependencies between classes, were
extracted with the tool Dependency Finder\footnote{Available at
\url{http://depfind.sf.net}.}.

\subsubsection{Non-software networks} We have collected 66 networks from distinct
domains, such as biology, sociology, technology, and linguistics. These networks
are freely available on the Internet and have been used in previous researches.
CITE

% S-score

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation of the Similarity Metric}

%In order to evaluate the similarity metric, we measured the similarity between
%the networks in the data set.

For the purposes of this research, a similarity metric must fullfil two
conditions: (i) it must yield high similarity between software networks, and
(ii) it must yield lower similarity between software networks and networks from
other domains.

Using the data set we can define S-score, a metric that represents how much a
particular network resemble software networks. It is defined as the average
similarity between the network and a sample of software networks:

$$
\mathrm{S\mbox{-}score}(a) = \frac{
\displaystyle\sum_{s \in S} \mathrm{similarity}(a, s)
}{|S|} \mathrm{,}
$$

where $S$ is the set of sample software networks, and $|S|$ is the number of
networks in $S$. In this work we use the full software data set consisting of 65
software networks as our sample.

We used the tool igraph \cite{igraph} to extract the TCP for each network in the
data set. Then, we measured the S-score for each software network, which ranged
from 0.83 to 0.98. The average S-score was 0.97 and the standard deviation,
0.03. The high average S-score and the low standard deviation show that the
metric successfully characterizes software networks by capturing their common
structural patterns.

%TODO: Boxplot by network class (software, metabolic, neural, social etc.)

Then we measured the S-score for each non-software network. The majority of the
networks (97.0\%) had a S-score lower than 0.83, which is the lowest S-score for
software networks in the sample. Some networks, e.g., the friendship networks
between students, showed negative S-score, meaning that they are very different
from software networks.

Two networks, though, showed high S-score, above 0.83. The network of links
between blogs on politics showed S-score 0.97. The neural network of the worm C.
Elegans also showed a high S-score (0.88). Further investigation is needed in
order to discover the reasons behind the high values and whether auxiliary
metrics can differentiate these networks from software networks.

\subsection{A Network Classification Model} \label{sec:classmodel}

Although the S-score of a network tells how close it is from software networks,
it does not tell whether a network is close enough that it can be considered
software-like. What is needed is a binary classification model that
distinguishes software-like networks from the other networks. The distinction
can be made by choosing a suitable S-score threshold. Networks with S-score
below the threshold are considered dissimilar from software networks; only
networks with S-score above the threshold are considered software-like.

As we have shown on the previous section, there are non-software networks with
high S-scores, hence it is impossible to build a perfect classification model,
regardless of the threshold. Nonetheless, a classification model can be
evaluated by its precision and recall. Consider our data set with both software
and non-software networks. Let $S$ be the set of all software networks, and $L$
the set of all networks that were classified by the model as software-like. The
precision of the model is

$$
\mathrm{precision}: ~\frac{S \cap L}{L},
$$

and the recall is

$$
\mathrm{recall}: ~\frac{S \cap L}{S}.
$$

Increasing the threshold has the effect of reducing the recall, because fewer
software networks are classified as software-like. Decreasing the threshold has
the effect of reducing the precision, since more non-software networks are
classified as software-like.

The choice of a proper threshold, thus, depends on whether it is more important
to have high precision or high recall. Because our research hypothesis is that
networks synthesized by the presented models are software-like, higher precision
means a stronger test, as fewer networks are classified as software-like.

To get 100\% precision, the threshold needs to be 0.98, so the non-software
network with highest S-score is below the threshold. The recall in this case,
though, would be too low, because most software networks would be misclassified.
So we chose the value 0.88, that is immediatly above the second greater S-score
for a non-software network. With this value, we have both high recall (95.4\%)
and high precision (96.9\%).

%threshold 0.88: 2/66 non-sw misclassified, 3/65 sw misclassified
%precision: 62 / 64
%recall: 62 / 65

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Evaluation of Network Models} \label{sec:evaluation}
% v.  differentiate, differ; distinguish; contrast; discern

In the previous section it was shown that many networks, although scale-free,
can be distinguished from software networks by a simple classification model
based on triad concentration profiles. In this section we show empirically that
the three network models previously presented synthesize networks that are
indistinguishable from software networks.
The experiment consists of synthesizing networks using many combinations of
parameters from the three models, and then classifying each network as
software-like or non software-like. %After that, we present rules for the choice
%of parameter values that lead to software-like networks.

\subsection{Synthetic Data Set}

We want to investigate if, with a proper choice of parameters, a model is
capable of synthesizing a network that resembles software networks. Because the
possible combinations of parameter values are infinite, we have set the number
of vertices to 1000 and then varied the remaining parameters in discrete steps.
In this section we describe the combinations of parameters values used for each
model.

\subsection{BCR+ networks}

We have chosen five different module graphs, which where extracted from actual
dependencies between Java archives distributed with five different software
systems of our sample: GEF (2 archives), iBATIS (4 archives), MegaMek (8
archives), findbugs (16 archives), and zk (32 archives). Because many of these
archives are intended to be reused in distinct projects, they provide a good
coarse-grained approximation of the concept of module.

%The module dependency networks are shown on Figure \ref{fig:architectures}.
%\begin{figure*}[!t] \centering
%\includegraphics[width=1.0\textwidth]{architectures} \caption{Module dependency
%graphs} \label{fig:architectures} \end{figure*}

For the remaining parameters, the following values were chosen:

\begin{itemize}
\item $p_1, p_2, p_3 \in \{0.0, 0.2, 0.4, 0.6, 0.8, 1.0\}$, with $p + q + r = 1$ and
$p + q > 0$;
\item $\din, \dout \in \{0, 1, 2, 3, 4\}$;
\item $\mu \in \{0.0, 0.2, 0.4, 0.6\}$.
\end{itemize}

We chose $p + q > 0$ because otherwise no node would be added after the
creation of the initial network. We also avoided large values for $\mu$ in
order to ignore networks with strongly coupled modules.

In total, 9,500 BCR+ networks were synthesized.

\subsection{CGW networks}

The parameters values were chosen as such:

\begin{itemize}
\item $p1, p2, p3, p4 \in \{0.0, 0.2, 0.4, 0.6, 0.8, 1.0\}$, with $p1 > 0$ and
$p1 + p2 + p3 + p4 = 1$;
\item $e1, e2, e3, e4 \in \{1, 2, 4, 8\}$;
\item $\alpha \in \{-1, 0, 1, 10, 100, 1000\}$;
\item number of modules: $m \in \{2, 4, 8, 16, 32\}$.
\end{itemize}

Notice that $e_i$ has no effect when $p_i = 0$; in this case $e_i$ was just set
to zero.

In total, 38,790 CGW networks were synthesized.

\subsection{LF networks}

The following values were chosen:

\begin{itemize}
  \item mixing parameter: $\mu \in \{0.0, 0.2, 0.4, 0.6\}$;
  \item degree exponent distribution: $\gamma \in \{2.18, 2.70, 3.35\}$;
  \item module size distribution: $\beta \in \{0.76, 0.99, 1.58\}$;
  \item average degree: $k \in \{5, 10, 15, 25\}$;
  \item maximum degree: $max_k \in \{58, 157, 482\}$;
  \item minimum module size: $min_m \in \{1, 10, 273\}$.
\end{itemize}

In order to choose these values, we analyzed software networks from our sample
with approximately 500 to 2,000 nodes, so no network was much bigger or much
smaller than the synthetic networks. We computed the exponents for the degree
and module size distributions using the maximum likelihood estimation method
\cite{Clauset2007}, and then chose the minimum, median and maximum values. For
$k$, $maxk$, and $min_m$, the values extracted from the software networks were
divided by the number of nodes in the network and then multiplied by 1000. We
then selected the minimum, median and maximum values. The parameter $max_m$ was
left unbound to avoid impossible combinations of parameters.

In total, 1,296 LF networks where synthesized.

\subsection{Results}
% This subsection needn't to be big. It's just numbers.

Each synthesized network was classified as software-like or non software-like,
using the classification model presented in section \ref{sec:classmodel}. The
results are summarized in Table \ref{tab:results}.

\begin{table}
\caption{Results for the classification of synthetic networks}
\centering
\begin{tabular}{|l|l|}
\hline
Model & Networks classified as software-like \\
\hline 
\hline
BCR+ & 21.18\% \\ % 2012 / 9500
\hline
CGW  & 19.40\% \\  % 7524 / 38790
\hline
LF   & 31.25\% \\ %  405 / 1296
\hline
\end{tabular}
\label{tab:results}
\end{table}

All models synthesized both software-like and non software-like networks. The
proportion of software-like networks was greater than 19\% for all models,
discarding the possibility that this result was obtained by pure chance.

%Show some graphs: histogram of average correlations for each model.
%\ref{fig:histograms}
%\begin{figure*}[!t]
%\center
%
%\subfigure[CGW]{\includegraphics[width=1.0in]{hist-cgw}}
%
%\subfigure[BCR+]{\includegraphics[width=1.0in]{hist-bcr}}
%
%\subfigure[LF]{\includegraphics[width=1.0in]{hist-lf}}
%
%\caption{Triad profiles}
%\label{fig:histograms}
%\end{figure*}

Of course, this result is of little practical value unless there is a
relationship between parameter values and S-score. For the purpose of this
research, it is important to know which values are more likely to lead to
software-like networks.

The algorithm 1R from data mining was used to help discover such relationship.
It analyzes the parameters and the classification of each network and finds a
rule that relates the value of one single parameter with the classification. We
are interested in rules for networks classified as software-like and in the
accuracy of these rules, i.e., the proportion of networks that are correctly
classified. The rules found by 1R are shown in Table \ref{tab:rules}.

\begin{table}
\caption{Rules for predicting the classification of a synthetic network. S
stands for software-like and N stands for non software like.}
\centering
\begin{tabular}{|l|l|l|}
\hline
Model & Rule & Accuracy \\
\hline 
\hline
\multirow{2}{*}{BCR+}
     & $p_1 \ge 0.7 \Rightarrow S$ & \multirow{2}{*}{82.4\%}  \\ 
     & $p_1 < 0.7 \Rightarrow N$ & \\ 
\hline
\multirow{2}{*}{CGW}
     & $p_1 \ge 0.5 \Rightarrow S$ & \multirow{2}{*}{82.3\%} \\  
     & $p_1 < 0.5 \Rightarrow N$ & \\  
\hline
\multirow{2}{*}{LF}   
     & $\gamma \ge 2.44 \Rightarrow S$ & \multirow{2}{*}{78.9\%} \\ 
     & $\gamma < 2.44 \Rightarrow N$ & \\ 
\hline
\end{tabular}
\label{tab:rules}
\end{table}

The rules are very simple and, thus, easy to follow. Despite their simplicity,
they have high accuracy, approximately 80\% in all models.  
%These rules allow one to know if a particular combination of parameter values
%is likely to lead to software-like networks.

%\subsection{Parameter-Based Classification Models}
%
%1R

%Naive Bayes

%\subsection{Homogeinity}
%
%Pick realistic networks from a model. Are they similar to each other? (see
%standard deviation) In other words: do the parameters make a difference?
%
%Are they similar to networks generated by the other models? In other words: are
%the models equivalent?

%\section{Threats to validity}
%
%% External validity (EV): can it be generalized?
%
%%Structural information isn't enough for a expert do produce a
%%decomposition (s/he may use data such as names and external documentation). 
%%
%%Even when considering only structural information, is it true
%%that experts would find a decomposition similar to the reference decomposition
%%imposed by the model?
%
%We generated only one network for each set of parameters. (as we've shown, some
%parameters are redundant as they do not change significantly the realism)
%
%Some clustering algorithms use weights and they weren't studied here.
%
%EV: We've only studied 65 systems, which is not that much.
%
%We only studied object-oriented systems implemented in Java. Maybe the results
%would be different if we studied systems implement in other languagens or using
%other paradigms. The choice of a particular technique for extracting
%dependencies (static analysis) may also have impact on the structure of the
%networks.
%
%% Koschke says that experts decompositions vary by no more than 80? percent.

\section{Conclusion and Future Work} \label{sec:conclusion}
% revised 2009-09-03

We have shown empirically that network models found in the literature can
synthesize networks that resemble the network of dependencies between classes in
object-oriented systems. This result supports the use of synthetic networks in
the evaluation of software clustering algorithms.
%that operate on class dependency networks.

The use of synthetic data is common in distributed computing research, but still
underexplored in software engineering research. Because many reverse engineering
tasks rely on dependency data \cite{Tonella2007}, we expect this work to have
impact beyond the software clustering community.

%Although the use of models and synthetic data is somewhat common in distributed
%computing research, it is underexplored in software engineering research. We
%expect this work to contribute to explore t
%common in distributed computing research, simulation is underexplored
%in software engineering. This work opens a door to the usage of simulation in
%the field of reverse engineering of software in order to evaluate RE 
%algorithms.

We accept that it is important to evaluate the algorithms with real software
networks, but we argue that the use of synthetic networks in a complementary
manner can give researchers new insights about the algorithms. First, the use of
models allows the creation of large test sets, thus diminishing the small sample
effects. Moreover, the networks are created in a controlled way, according to
model parameters, so it is possible to study the behavior of the algorithms with
different parameter values.

In a future work, we intend to use synthetic networks in the evaluation of
software clustering algorithms that were previously studied with real networks
\cite{Wu2005}. After that we will be able to compare the results obtained by
the approaches.

\appendix %{Network Data Set}

This appendix lists the networks used in this work.

\subsection{Software Networks}

Systems hosted by SourceForge (\url{http://sourceforge.net/}):
\begin{itemize}
\item AbaGuiBuilder-1.8
\item alfresco-labs-deployment-3Stable
\item aoi272
\item stendhal-0.74
\item battlefieldjava-0.1
\item checkstyle-5.0
\item dom4j-1.6.1
\item findbugs-1.3.8
\item freetts-1.2.2-bin
\item ganttproject-2.0.9
\item geoserver-2.0-beta1-bin
\item geotools-2.5.5-bin
\item gfp\_0.8.1
\item hibernate-distribution-3.3.1.GA-dist
\item hsqldb\_1\_8\_0\_10
\item iBATIS\_DBL-2.1.5.582
\item iReport-nb-3.5.1
\item JabRef-2.5b2-src
\item jailer\_2.9.9
\item jalopy-1.5rc3
\item jasperreports-3.5.2-project
\item jfreechart-1.0.13
\item pentaho-reporting-engine-classic-0.8.9.11
\item jGnash-2.2.0
\item jgraphpad-5.10.0.2
\item jmsn-0.9.9b2
\item juel-2.1.2
\item JXv3.2rc2deploy
\item makagiga-3.4
\item MegaMek-v0.34.3
\item iFreeBudget-2.0.9
\item mondrian-3.1.1.12687
\item oddjob-0.26.0
\item openxava-3.1.2
\item pdfsam-1.1.3-out
\item pjirc\_2\_2\_1\_bin
\item pmd-bin-4.2.5
\item proguard4.3
\item smc\_6\_0\_0
\item squirrel-sql-3.0.1-base
\item squirrel-sql-3.0.1-standard
\item tvbrowser-2.7.3-bin
\item villonanny-2.3.0.b02.bin
\item rapidminer-4.4-community
\item zk-bin-3.6.1
\end{itemize}

Systems hosted in other sites:

\begin{itemize}
\item ArgoUML-0.28
\item GEF-0.13-bin
\item Hl7Comm.1.0.1
\item IRPF2009v1.1
\item broker-4.1.5 (OurGrid)
\item dbwrench
\item ec2-api-tools
\item ermodeller-1.9.2-binary
\item flyingsaucer-R8
\item gdata-src.java-1.31.1
\item guice-2.0
\item gwt-windows-1.6.4
\item jai-1\_1\_4-pre-dr-b03-lib-linux-i586-08\_Jun\_2009
\item jakarta-tomcat-5.0.28-embed
\item juxy-0.8
\item myjgui\_0.6.6
\item peer-4.1.5 (OurGrid)
\item subethasmtp-3.1
\item thinkui\_sqlclient-1.1.2
\item worker-4.1.5 (OurGrid)
\end{itemize}

\subsection{Networks from Other Domains}

\begin{itemize}
\item 5 friendship networks from Facebook \cite{Traud2008}
\item 3 electronic circuit networks \cite{Milo2004}
\item 4 word adjacency networks \cite{Milo2004}
\item 3 protein structure networks \cite{Milo2004}
\item 2 social networks of positive sentiment \cite{Milo2004}
\item 43 metabolic networks \cite{Jeong2000}
\item Protein interaction network for yeast \cite{Jeong2001}
\item Links between political blogs \cite{Adamic2005}
\item Neural network of C Elegans \cite{Watts1998}
\item Network ``beta3sreduced'' (unknown source)
\item Network ``czech'' (unknown source)
\item Network ``ecoli-metabolic'' (unknown source).
\end{itemize}