### 2010-01-05 trunk/latex/dissertacao/conclusao.tex
Adicionando template e estrutura da dissertação de mestrado.

% 2-10 páginas

%   (Limitações)
%   (Lições aprendidas)
% Recapitulação
% Resumo das contribuições
% Trabalhos futuros

\chapter{Conclusão}