### 2010-01-05 trunk/latex/dissertacao/estudodecaso.tex

% (5-20)

\chapter{Estudo de Caso: Comparando Algoritmos de Clustering}